---
title: 'Assignment 1: Sierra Amphibians'
author: "Anastasia Kunz"
date: "1/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(readxl)
library(lubridate)
library(patchwork)
```



```{r}
# Read in data
frog_data <- read_excel(here("data","sierra_amphibians.xlsx"))
  
```
#### Total Mountain Yellow-Legged Frog count each year

```{r}

total_yellow <- frog_data %>% 
  filter(amphibian_species == "RAMU") %>% 
  filter(amphibian_life_stage != "EggMass") %>% 
  mutate(date = ymd(survey_date)) %>% 
  mutate(year = year(date)) %>% 
  drop_na(amphibian_number) %>% 
  count(year, amphibian_life_stage, wt = amphibian_number)


year_totals <-ggplot(total_yellow, aes(x = year, y = n))+
  geom_col(fill = "darkslategray3")+
  facet_wrap(~amphibian_life_stage)+
  labs(
    x = "Year",
    y = "Total Count",
    title = "Mountian Yellow-Legged Frog Count by Life Stage"
  )
  
# Question: not sure best way to present data with the different life stages having such spread out values; when displaying tadpoles it makes it hard to read adult stage. Would the stacked graph be better? using fill = amphibian_life_stage to show the relative amounts? or is that harder to read?


```


#### Top 5 lakes with greatest Adult and SubAdult counts of Yellow-Legged Frogs
```{r}

lake_counts <- frog_data %>% 
  filter(amphibian_species == "RAMU") %>% 
  filter(amphibian_life_stage == c("Adult", "SubAdult")) %>% 
  mutate(lake_id = as.character(lake_id)) %>% 
  mutate(lake_id = paste("Lake", lake_id, sep = " ")) %>% 
  count(lake_id, wt = amphibian_number) %>% 
  rename(total_amphibian = n) %>% 
  slice_max(total_amphibian, n = 5)

top5 <- ggplot(lake_counts, aes(x = fct_reorder(lake_id, total_amphibian), y = total_amphibian))+
  geom_col(fill = "darkslategray3")+
  coord_flip()+
  labs(
    x = "Lake ID",
    y = "Total Adult and SubAdult Count",
    title = "Top 5 Lakes with Highest Mountain Yellow-Legged Frogs",
    subtitle = "Adult and SubAdult Rana muscosa, 1995 - 2002"
  )

```

#### Data Visualization for Mountain Yellow-Legged Frogs

```{r, fig.cap = "**Figure 1**. Mass (g) vs height (m)"}

patchwork_graph <- (year_totals/top5) & theme_minimal()

patchwork_graph
```

#### References

Knapp, R.A., C. Pavelka, E.E. Hegeman, and T.C. Smith. 2020. The Sierra Lakes Inventory Project: Non-Native fish and community composition of lakes and ponds in the Sierra Nevada, California ver 2. Environmental Data Initiative. https://doi.org/10.6073/pasta/d835832d7fd00d9e4466e44eea87fab3




